Surface-driven registration method for the structure-informed segmentation of diffusion MR images\tnoteref{t1}

### 2015-11-04 Publications/2015-NeuroImage/04-methods.tex
add final MJLC comments

\begin{figure*}
\includestandalone[width=\linewidth]{figure02-model}
% \includegraphics[width=\linewidth]{figure04A}
% \includegraphics[width=\linewidth]{figure04B}
% \includegraphics[width=\linewidth]{figure04C}
\caption{%
\revcomment{%
Evolution of the segmentation model defined by the homogeneous regions $\Omega_l$
  for a real dataset.
A) shows the initialization of the model, when the anatomically-correct surfaces are not
  mapped into the \gls*{dmri} space.
The panel shows the resulting conditional distributions of segmenting the (distorted) \gls*{fa} and
  \gls*{adc} maps with the original (undistorted) surfaces.
Panel B shows the equivalent information, using the surfaces with perfect matching with the distorted space
  of \gls*{dmri}.
Since the surfaces are now aligned in $M$, the spreads of the conditional distributions for each region are
  reduced with respect to those in A.
Panel C, after \regseg{} registration, shows how the covariance matrices $\boldsymbol{\Sigma}_l$
  are minimized with respect to panel A, and the resulting segmentation is close
  to the theoretical situation presented in panel B.
The left (large) plot in each panel represents a kernel density estimate of the location and spread of
  each region $\Omega_l$ in the bivariate feature space of $M$, which comprises
  the \gls*{fa} and \gls*{adc} maps.
At the top and the right margins, the marginal distributions of $\Omega_l$ are also plotted for
  the \gls*{fa} and \gls*{adc}, respectively.
On the right, the sampled distribution of each region is shown in small scatter plot, in contrast to
  the complete bivariate data.}
}\label{fig:regseg-model}
\todo[inline]{C5 (R.1)}
\end{figure*}

>>>>

\begin{figure*}
\includestandalone[width=\linewidth]{figure02-model}
% \includegraphics[width=\linewidth]{figure04A}
% \includegraphics[width=\linewidth]{figure04B}
% \includegraphics[width=\linewidth]{figure04C}
\caption{%
\revcomment{%
Evolution of the segmentation model defined by the homogeneous regions $\Omega_l$
  for a real dataset.
The left plot in each panel represents a kernel density estimate of the location and spread of
  each region $\Omega_l$ in the bivariate feature space of $M$, which comprises
  the \gls*{fa} and \gls*{adc} maps.
At the top and the right margins, the marginal distributions of $\Omega_l$ on the \gls*{fa} and
  \gls*{adc} are respectively displayed.
On the right side, the sampled distribution of each region is shown in small scatter plot, in contrast to
  the complete bivariate data.
A) shows the initialization of the model, when the anatomically-correct surfaces are not
  mapped into the \gls*{dmri} space.
The panel shows the resulting conditional distributions of segmenting the (distorted) \gls*{fa} and
  \gls*{adc} maps with the original (undistorted) surfaces.
Panel B shows the equivalent information, using the surfaces that are perfectly mapped into the
  distorted \gls*{dmri} space.
Since the surfaces are now aligned in $M$, the spreads of the conditional distributions for each region are
  reduced with respect to those in A.
Panel C, after \regseg{} registration, shows how the covariance matrices $\boldsymbol{\Sigma}_l$
  are minimized with respect to panel A, and the resulting segmentation is close
  to the theoretical situation presented in panel B.}
}\label{fig:regseg-model}
\todo[inline]{C5 (R.1)}
\end{figure*}

### 2015-11-04 Publications/2015-NeuroImage/06-discussion.tex
add final MJLC comments

% -*- root: 00-main.tex -*-
\section{Discussion}
\label{sec:regseg-discussion}
\revcomment[C10 (R.2)]{%
We present \regseg{}, a simultaneous segmentation and registration method that
  maps a set of nested surfaces into a multivariate target-image.}
\revcomment[C1 (R.1)]{%
The nonlinear registration process evolves driven by the fitness of the
  picewise-smooth classification of voxels in the target volume imposed
  by the current mapping of the surfaces.}
\revcomment[moved]{%
We propose \regseg{} to map anatomical information extracted from \gls*{t1}
  images into the corresponding \gls*{dmri} of the same subject.
Previously, joint segmentation and registration has been applied successfully to
  other problems such as longitudinal object tracking \citep{paragios_level_2003}
  and atlas-based segmentation \citep{gorthi_active_2011}.
The most common approach involves optimizing a deformation model (registration)
  that supports the evolution of the active contours (segmentation),
  like \cite{paragios_level_2003,yezzi_variational_2003}.}
% Conversely, in structure-informed segmentation, the sources of variability are the geometrical distortions
%   after imaging and the anatomical evolution in longitudinal studies.}
\revcomment[C3 (R.1)]{%
\Regseg{} can be seen as a particular case of atlas-based segmentation-registration methods,
  replacing the atlas by the structural image of the subject (\emph{structure-informed segmentation}).
The main difference of atlas-based segmentation and the application at hand is the resolution of the
  target image.
Atlas-based segmentation is typically applied on high-resolution images.
A comprehensive review of joint segmentation and registration methods applied in atlas-based
  segmentation is found in \citep{gorthi_active_2011}.
They also propose a multiphase level-set function derived from a labeled atlas to implement the
  active contours that drive the atlas registration.
Alternatively, \regseg{} implements the active contours with a hierarchical set of explicit
  surfaces (triangular meshes) instead of the multiphase level sets, and registration
  is driven by shape-gradients \citep{herbulot_segmentation_2006}.
The use of explicit surfaces enables segmenting \gls*{dmri} images below pixel resolution.}

@@

\revcomment[C9 (R.2)]{%
The proposed application of the method in the task of identifying structural information
  in \gls*{dmri} images is an active field of research \citep{jeurissen_tissuetype_2015}.
Current processing steps involved in the connectome extraction require a precise segmentation
  of anatomical structures in the diffusion space.
Some examples of these processing tasks are the structure-informed reconstruction of \gls*{dmri}
  data \citep{jeurissen_multitissue_2014,daducci_accelerated_2015}, the anatomically constrained
  tractography \citep{smith_anatomicallyconstrained_2012}, and the imposition of the cortical
  parcellation mapped from the \gls*{t1} image \citep{hagmann_mapping_2008}.
The problem was firstly addressed using image segmentation approaches in the native diffusion
  space, without definite and compelling results.}
\revcomment[C18 (R.2)]{%
With the introduction of retrospective correction methods for the \emph{\gls*{epi} distortions}
  and image registration approaches, the task has been typically solved in a two-step approach.
First, the \glspl*{dwi} are corrected for \emph{\gls*{epi} distortions} by estimating
  the nonlinear-deformation field from extra MR acquisitions
  \citep{jezzard_correction_1995,chiou_simple_2000,cordes_geometric_2000,kybic_unwarping_2000}.
Second, mapping the structural information from the corresponding \gls*{t1} image
  using a linear registration tool like \emph{bbregister} \citep{greve_accurate_2009}.
The current activity on improving correction methods \citep{irfanoglu_drbuddi_2015} and
  the comeback of segmentation of \gls*{dmri} in its native space
  \citep{jeurissen_tissuetype_2015} proof the open interest of this application.
\Regseg{} addresses this joint problem in a single step and it does not require any additional
  acquisition other than the minimal protocol comprehending only \gls*{t1} and \gls*{dmri} images.
This situation is found commonly in historical datasets.
Moreover, since the structural information is projected into the native space of \gls*{dmri},
  the whole connectome extraction process can be performed without resampling data to an
  undistorted space.}

>>>>

% -*- root: 00-main.tex -*-
\section{Discussion}
\label{sec:regseg-discussion}
\revcomment[C10 (R.2)]{%
We present \regseg{}, a simultaneous segmentation and registration method that
  maps a set of nested surfaces into a multivariate target-image.}
\revcomment[C1 (R.1)]{%
The nonlinear registration process evolves driven by the fitness of the
  picewise-smooth classification of voxels in the target volume imposed
  by the current mapping of the surfaces.}
\revcomment[moved]{%
We propose \regseg{} to map anatomical information extracted from \gls*{t1}
  images into the corresponding \gls*{dmri} of the same subject.
Previously, joint segmentation and registration has been applied successfully to
  other problems such as longitudinal object tracking \citep{paragios_level_2003}
  and atlas-based segmentation \citep{gorthi_active_2011}.
The most common approach involves optimizing a deformation model (registration)
  that supports the evolution of the active contours (segmentation),
  like \cite{paragios_level_2003,yezzi_variational_2003}.}
% Conversely, in structure-informed segmentation, the sources of variability are the geometrical distortions
%   after imaging and the anatomical evolution in longitudinal studies.}
\revcomment[C3 (R.1)]{%
\Regseg{} can be seen as a particular case of atlas-based segmentation-registration methods,
  replacing the atlas by the structural image of the subject (\emph{structure-informed segmentation}).
The main difference of atlas-based segmentation and the application at hand is the resolution of the
  target image.
Atlas-based segmentation is typically applied on structural and high-resolution images.
A comprehensive review of joint segmentation and registration methods applied in atlas-based
  segmentation is found in \citep{gorthi_active_2011}.
They also propose a multiphase level-set function initialized from a labeled atlas to implement
  the active contours that drive the atlas registration.
Alternatively, \regseg{} implements the active contours with a hierarchical set of explicit
  surfaces (triangular meshes) instead of the multiphase level sets, and registration
  is driven by shape-gradients \citep{herbulot_segmentation_2006}.
As an advantage, the use of explicit surfaces enables segmenting \gls*{dmri} images
  with accuracy below pixel size.}

@@

\revcomment[C9 (R.2)]{%
The proposed application of the method in the task of identifying structural information
  in \gls*{dmri} images is an active field of research \citep{jeurissen_tissuetype_2015}.
Current processing steps involved in the connectome extraction require a precise segmentation
  of anatomical structures in the diffusion space.
Some examples of these processing tasks are the structure-informed reconstruction of \gls*{dmri}
  data \citep{jeurissen_multitissue_2014,daducci_accelerated_2015}, the anatomically constrained
  tractography \citep{smith_anatomicallyconstrained_2012}, and the imposition of the cortical
  parcellation mapped from the \gls*{t1} image \citep{hagmann_mapping_2008}.
The problem was firstly addressed using image segmentation approaches in the native diffusion
  space, without definite and compelling results.}
\revcomment[C18 (R.2)]{%
With the introduction of retrospective correction methods for the \emph{\gls*{epi} distortions}
  and image registration approaches, the task has been typically solved in a two-step approach.
First, the \glspl*{dwi} are corrected for \emph{\gls*{epi} distortions} by estimating
  the nonlinear-deformation field from extra MR acquisitions
  \citep{jezzard_correction_1995,chiou_simple_2000,cordes_geometric_2000,kybic_unwarping_2000}.
Second, mapping the structural information from the corresponding \gls*{t1} image
  using a linear registration tool like \emph{bbregister} \citep{greve_accurate_2009}.
The current activity on improving correction methods \citep{irfanoglu_drbuddi_2015} and
  the comeback of segmentation of \gls*{dmri} in its native space
  \citep{jeurissen_tissuetype_2015} proof the open interest of this application.
\Regseg{} addresses this joint problem in a single step and it does not require any additional
  acquisition other than the minimal protocol comprehending only \gls*{t1} and \gls*{dmri} images.
This situation is found commonly in historical datasets.
Moreover, since the structural information is projected into the native space of \gls*{dmri},
  the whole connectome extraction process can be performed on the unaltered \gls*{dmri}
  signal (i.e. without resampling data to an undistorted space).}
